---
title: "Notes on R package development"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prerequisites

Package development is based on the helper package `usethis`, and 
`here` package to set the project root. The notes are based on the information
provided in [**this package online reference**](http://usethis.r-lib.org/)

```{r}
library("usethis")
library("here")
```

1. Create a folder where you want to start the self-contained package
2. Navigate to this folder and open a terminal from this folder
3. Start RStudio from the terminal
```{bash, eval=FALSE}
rstudio
```

After this, RStudio opens with the working directory in the package folder.

# Set up the package root manually

Next, you can set the project root programatically, for example using the `here`
package. You can also set up a package project using the 
`File` -> `New Project...` menu in RStudio.

To initate a hidden `.here` file that sets the project root within the 
self-contained project folder, you can use:

```{r, eval=FALSE}
here::set_here()
```

After setting the project root, you can store the active project for use
for the remaining session. For details, consider
[**this help section**](http://usethis.r-lib.org/reference/proj_get.html)

To create the package template folder structure and key files you can
run:

```{r, eval=FALSE}
create_package(here())
```

This creates the package template and starts the RStudio project session in a
new session.


# Use `usethis` interactively

[**This `r-lib` online article**](http://usethis.r-lib.org/articles/articles/usethis-setup.html)
explains further how to setup and invoke R package development workflow
functions. It provides a good overview how to integrate git version control and
how to install package building tools, e.g. compilers.

1. Set up interactive mode. This will open the `.Rprofile` file 

```{r, eval=FALSE}
usethis::use_usethis()
```
2. Copy the following lines into the file:
```{r, eval=FALSE}
if (interactive()) {
  suppressMessages(require(devtools))
  suppressMessages(require(usethis))
}
```



# Initiate a github repositiory 


